---
title: "Guide to tables"
output:
  html_document: default
  html_notebook: default
  word_document: default
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

##  libraries
library(tidyverse)
library(stargazer)
library(MatchIt)
library(Matching)
library(htmlTable)
```

##  Introduction

We want to create journal quality tables that can be 
1) exported to excel
2) easy modifiable within Rmarkdown (or R)

There are many ways to write tables using Rmarkdown, LaTex, and html. However,
the code for doing so can get very complicated. Furthermore, many tables require
pre-processing (i.e. model summaries, descriptive summaries) that is labourious
to write by hand.

Therefore I have complied examples of my go-to code for writing different table
types. I want to basically embed the R code for making tables in Rmarkdown and
knit the document as html for copy and pasting tables into word. Most of these
pacakges will output html or rmarkdown code directly so it is important to remember:

1) Whether they are outputting html or markdown for further processing
2) Use results = 'asis' in the r chunks to stop applying further markup on top of the html

##  Reading in data and tables

This just reads in the example data that I will use for making tables.

```{r read and clean, results = 'hide'}
table <- 
  read.csv('Example data/Example RCI table.csv', 
           check.names = F # This option stops R from replacing space with . in the col header
           ) 

data <- 
  read.csv('Example data/Example RCI data.csv', 
           check.names = F # This option stops R from replacing space with . in the col header
           ) 

```
##  1) Directly output R object as formatted table (no pre-processing)

This is an example where we simply load in an existing table (or dataframe). We
then directly out the results as a html table to be written up. We use htmlTable
but another package such as xtable or so can do the trick. htmlTable is still
earlish in development. Other ways to do this include ktable. 

see: https://cran.r-project.org/web/packages/htmlTable/vignettes/tables.html
```{r basic stargazer table, results = 'asis'}
col.names <- c('', 
               'Poor', 
               'Non-poor')
cgroup <- c('Zones ranked by <br> distance to centre&dagger;',
            'Cumulative proportion of:')

table %>% htmlTable(
  header = col.names,
  cgroup = cgroup,
  n.cgroup = c(1, 2),
  rnames = F,  #No row names
  caption = 'Example of table used for Duncan Index',
  tfoot = '&dagger; Ranked from nearest to furthest'
  )  
```

##  2a) Creating descriptive tables

For more descriptive tables, we want a package to easily do the summaries AND output
the tables we desire. We can use arsenal to preprocess the data and write it in rmarkdown. 

see: https://cran.r-project.org/web/packages/arsenal/vignettes/tableby.html#pretty-rmarkdown-version-of-table

```{r, results = 'asis'}
library(arsenal)

tab.obj <- tableby(~., data
                   #weights = ,
                   )
labels(tab.obj) # can change these


summary_table <- arsenal:::summary.tableby(
  tab.obj,
  labelTranslations = list(c.propA = 'Proportion in A',
  c.propB = 'Proportion in B'),
  title = 'Example',
  pfootnote = T #show footnote for p values test
  )
```
Footnote: Adding footnotes in arsenal is just using rmd below the table https://github.com/eheinzen/arsenal/issues/171
Since arsenal seems to do alot of the work in rmarkdown

##  2b) Multiple summaries / subsets
```{r, results = 'asis'}
dataMore <- 
  rbind(data, data)
dataMore$sample <- 1:2 %>% as.character


tab.obj <- tableby(sample ~., dataMore
                   #weights = ,
                   )
labels(tab.obj) # can change these


arsenal:::summary.tableby(tab.obj,
                          labelTranslations = list(c.propA = 'Proportion in A',
                                                   c.propB = 'Proportion in B'),
                          title = 'Example',
                          pfootnote = T #show footnote for p values test
                          ) 
```
